Capstone Movielens Report

### 2025-12-12 reports/capstone-movielens.site/05.appendix.Rmd
Function described: `calc_UMGY_SmoothedDay_effect.MSE`.

::: {.noteblock data-latex=""}
This helper function is used internally in the [calc_UMGY_effect_MSE.cv](#func.calc_UMGY_effect_MSE.cv) function (described below in this section) as shown in the [Usage](#func.calc_UMGY_effect_MSE.usage) subsection of *This Function's Description Section* below.
:::

@@

The function calculates *Mean Squared Error (MSE)* for the *UMGYE Model* passed in the [UMGY_effect](#func.calc_UMGY_effect_MSE.args.UMGY_effect) argument using the validation dataset passed in the [test_set](#func.calc_UMGY_effect_MSE.args.test_set) arg. 
:::

@@

This function is also called within other helper functions, such as the [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) and [regularize.train_UMGYD_effect.cv](#func.regularize.train_UMGYD_effect.cv) (described below in sections [UMGYDE Model: Tuning `loess` Params] and [UMGYDE Model: Regularization] of *this Appendix*, respectively). The usage in the latter function is shown in the [Usage](#func.train_UMGY_SmoothedDay_effect.cv.usage) section of *this Description* below.

@@

Predicts *user rating* values on the *test dataset* passed in the [test_set](#func.UMGY_SmoothedDay_effect.predict.args.test_set) argument based on the data of the trained *UMGY Effect (UMGYDE) Model* passed in the [day_smoothed_effect](#func.UMGY_SmoothedDay_effect.predict.args.day_smoothed_effect) argument.

::: {.noteblock data-latex=""}
This function is also called within other helper functions, such as the [calc_UMGY_SmoothedDay_effect.MSE](#func.calc_UMGY_SmoothedDay_effect.MSE) and [UMGYDE_model.predict](#func.UMGYDE_model.predict) described below in this section and [Matrix Factorization Functions] section of *this Appendix*, respectively.
:::

@@

This function implements an algorithm for computing the *user rating values* in accordance with the *mathematical description* of the *UMGYDE Model*, as described in section [Mathematical Description of the UMGYDE Model] of *this Report*.
:::

@@

Validates the *UMGY Effect (UMGYDE) Model* produced by the  [train_UMGY_effect](#func.train_UMGY_effect) or [train_UMGY_effect.cv](#func.train_UMGY_effect.cv) function call.

::: {.noteblock data-latex=""}
This helper function is used internally in the [calc_UMGY_SmoothedDay_effect.MSE.cv](#func.calc_UMGY_SmoothedDay_effect.MSE.cv) function (described below in this section) as shown in the [Usage](#func.calc_UMGY_SmoothedDay_effect.MSE.usage) subsection of *This Function's Description Section* below.
:::

@@

@@

:::
::: { #func.calc_UMGY_SmoothedDay_effect.MSE.args.UMGYD_effect .sidebar }
  - **UMGYD_effect:** Data frame representing the *UMGYDE Model*.

@@

The function calculates *Mean Squared Error (MSE)* for the *UMGYDE Model* passed in the [UMGYD_effect](#func.calc_UMGY_SmoothedDay_effect.MSE.args.UMGYD_effect) argument using the validation dataset passed in the [test_set](#func.calc_UMGY_SmoothedDay_effect.MSE.args.test_set) arg. 
:::

@@

The source code of the [calc_UMGY_SmoothedDay_effect.MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L141) function is shown below:
```{r eval=FALSE}
```

@@

Validates the *UMGY Effect (UMGYDE) Model* (produced by the [train_UMGY_effect](#func.train_UMGY_effect) or [train_UMGY_effect.cv](#func.train_UMGY_effect.cv) function call) with use of *K-Fold Cross-Validation* as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV].

>>>>

@@

The function calculates *Mean Squared Error (MSE)* for the *UMGYE Model* passed in the [UMGY_effect](#func.calc_UMGY_effect_MSE.args.UMGY_effect) argument using the validation dataset passed in the [test_set](#func.calc_UMGY_effect_MSE.args.test_set) arg.

::: {.noteblock data-latex=""}
This auxiliary function is used internally in the [calc_UMGY_effect_MSE.cv](#func.calc_UMGY_effect_MSE.cv) function (described below in this section) as shown in the [Usage](#func.calc_UMGY_effect_MSE.usage) subsection of *this Description* above.
:::
:::

@@

::: {.noteblock data-latex=""}
This function is also called within other helper functions, such as the [train_UMGY_SmoothedDay_effect.RMSE.cv](#func.train_UMGY_SmoothedDay_effect.RMSE.cv) and [regularize.train_UMGYD_effect.cv](#func.regularize.train_UMGYD_effect.cv) (described below in sections [UMGYDE Model: Tuning `loess` Params] and [UMGYDE Model: Regularization] of *this Appendix*, respectively). The usage in the latter function is shown in the [Usage](#func.train_UMGY_SmoothedDay_effect.cv.usage) section of *this Description* below.
:::

@@

Predicts *user rating* values on the *test dataset* passed in the [test_set](#func.UMGY_SmoothedDay_effect.predict.args.test_set) argument based on the data of the trained *UMGYD Effect (UMGYDE) Model* passed in the [day_smoothed_effect](#func.UMGY_SmoothedDay_effect.predict.args.day_smoothed_effect) argument.

@@

This function implements an algorithm for computing the *user rating values* in accordance with the *mathematical description* of the *UMGYDE Model*, as described in section [Mathematical Description of the UMGYDE Model] of *this Report*.

::: {.noteblock data-latex=""}
This function is also called within other helper functions, such as the [calc_UMGY_SmoothedDay_effect.MSE](#func.calc_UMGY_SmoothedDay_effect.MSE) and [UMGYDE_model.predict](#func.UMGYDE_model.predict) described below in this section and [Matrix Factorization Functions] section of *this Appendix*, respectively.
:::
:::

@@

Validates the *UMGYD Effect (UMGYDE) Model* using data passed in the [day_smoothed_effect](#func.calc_UMGY_SmoothedDay_effect.MSE.args.day_smoothed_effect) argument on the *test dataset* passed in the [test_set](#func.calc_UMGY_SmoothedDay_effect.MSE.args.test_set) argument.

@@

```{r eval=FALSE}
    cv_fold_dat$validation_set |> 
      calc_UMGY_SmoothedDay_effect.MSE(day_smoothed_effect)
```

@@

:::
::: { #func.calc_UMGY_SmoothedDay_effect.MSE.args.day_smoothed_effect .sidebar }
  - **day_smoothed_effect:** Dataframe object representing the *UMGYDE Model*.

@@

This function commits the *UMGYDE Model* validation by calculating the *Mean Squared Error (MSE)*.

In *this Project*, the *UMGYDE* data object representing the *UMGYDE Model* (and being passed in the [day_smoothed_effect](#func.loess_de.args.de_bias.dat) argument) is created and returned by the following functions described above in *this Section*:
  - [calc_UMGY_SmoothedDay_effect](#func.calc_UMGY_SmoothedDay_effect) 
  - [train_UMGY_SmoothedDay_effect](#func.train_UMGY_SmoothedDay_effect) 
  - [train_UMGY_SmoothedDay_effect.cv](#func.train_UMGY_SmoothedDay_effect.cv).

Internally, this function calls the auxiliary functions [UMGY_SmoothedDay_effect.predict](#func.UMGY_SmoothedDay_effect.predict) (described above in this section), passing to it the value of the argument [day_smoothed_effect](#func.calc_UMGY_SmoothedDay_effect.MSE.args.day_smoothed_effect) in the same-named argument.

::: {.noteblock data-latex=""}
This auxiliary function is used internally in the [calc_UMGY_SmoothedDay_effect.MSE.cv](#func.calc_UMGY_SmoothedDay_effect.MSE.cv) function (described below in this section) as shown in the [Usage](#func.calc_UMGY_SmoothedDay_effect.MSE.usage) section of *This Description* above.
:::
:::

@@

The source code of the [calc_UMGY_SmoothedDay_effect.MSE](https://github.com/AzKurban-edX-DS/Capstone-MovieLens/blob/main/r/src/support-functions/UMGYD-effect.functions.R#L141) function is shown below:
```{r eval=FALSE}
calc_UMGY_SmoothedDay_effect.MSE <- function(test_set, day_smoothed_effect) {
  test_set |>
    UMGY_SmoothedDay_effect.predict(day_smoothed_effect) |>
    mutate(resid = rating - predicted) |> 
    pull(resid) |> mse()
}
```

@@

Validates the *UMGYD Effect (UMGYDE) Model* (produced by the [train_UMGY_effect](#func.train_UMGY_effect) or [train_UMGY_effect.cv](#func.train_UMGY_effect.cv) function call) with use of *K-Fold Cross-Validation* as described in [Section 29.6 *Cross validation / K-fold cross validation*](https://rafalab.dfci.harvard.edu/dsbook-part-2/ml/resampling-methods.html#k-fold-cross-validation) of the _Course Textbook (New Edition)_[@IDS2_29-6_KFCV].
